---
title: "LHneurons"
output: html_document
date: "2023-11-10"
---

look for all LH neurons, also those currently not annotated as LH
1. currently typed LH neurons
  including LHCENT, LHLN, ALIN
2. all neurons with more than 30% input form the LH and are in our meta data (meaning they are neurons)
  remove cell classes that we are not currently interested in
3. merge the two and use for LH to DN connectivity 

```{r setup, include=FALSE}
library(dplyr)
library(coconatfly)
library(tibble)
library(fafbseg)
library(arrow)
library(ggplot2)
```

1. typed neurons that include LH in their name
```{r setup, include=FALSE}
LH = cf_meta(cf_ids(flywire='/type:LH.+')) 
nrow(LH)
# 771, not all need to increase by looking at everything with more than x% input from the LH
table(LH$type)
# get all that have a cell class LHLN

# get all LH that have a LH hemibrain match and are currently a number in type
LHhemi = cf_meta(cf_ids(flywire='/hemibrain_type:LH.+')) 
difference2 = subset(LHhemi,!LHhemi$id %in% LH$id)
difference3   <-difference2[!is.na(as.numeric(difference2$type)), ]
LH = bind_rows(LH,difference3)


LHLN = cf_meta(cf_ids(flywire='/cell_class:LHLN')) 
nrow(LHLN)
LHLN <- LHLN %>%
  mutate(type = ifelse(grepl("^\\d+$", type), paste0("LHLN", type), type))
LHneurons = unique(bind_rows(LH, LHLN))
# give all without cell type an unknown so I can count it
LHneurons <- LHneurons %>%
  mutate_all(~ifelse(is.na(.), "unknown", .))
# 1223
# this is closer to what we expect for LH neurons
table(LHneurons$cell_class)
#      ALIN  LHCENT    LHLN unknown 
      # 2      42     479     700 

saveRDS(LHneurons, file = "LHtypedneurons_meta.RDS")
# start with these and then include the other LH relevant neurons later
```


2. neurons with postsynapses in the LH
# read from feather file with postsynapses per neuropil 
```{r setup, include=FALSE}
f="/Users/tomke/Library/Application Support/R/fafbseg/flywire_connectome_analysis_data/756/per_neuron_neuropilv2_filtered_count_post_756.feather"
ff = read_feather(f)
```

# tansform into percent
```{r innervating LH, echo=FALSE}
# transform into percent
ff$post_pt_root_id = as.character(ff$post_pt_root_id)
ffsum = aggregate(ff$count, by = list(ff$post_pt_root_id), FUN = sum)
names(ffsum) = c("post_pt_root_id", "sum")
ff = left_join(ff, ffsum)
ff$percent = (ff$count/ff$sum)*100
```

# how many have >60% postsynapses in the LH: 
```{r innervating LH, echo=FALSE}
LHff = subset(ff, ff$neuropil %in% c("LH_L", "LH_R"))
# take all with more than 5 synapses
LHff = subset(LHff, LHff$count>5)
# take all with more than 60%
LHff = subset(LHff, LHff$percent>60)
#update ids
LHff$post_pt_root_id = flywire_updateids(LHff$post_pt_root_id)
# still a lot of neurons or fragments
# chose only the ones that have some sort of meta annotation
flytable_login()
m =  flytable_meta()
LHff_m = subset(m, m$root_id %in% LHff$post_pt_root_id)
LHff_m <- LHff_m %>%
  mutate_all(~ifelse(is.na(.), "unknown", .))
table(LHff_m$cell_class)
         # ALIN          ALPN            AN           DAN        LHCENT          LHLN 
         #    4            97             1             2            15           479 
         #  mAL putative_glia       unknown 
         #    1            71          2425 

# remove cell classes that we are not interested in at the moment
LHff_m = subset(LHff_m, !LHff_m$cell_class %in% c("ALPN","AN", "DAN","putative_glia", "mAL" ))
LHff_mids = LHff_m$root_id
# 2646
   # ALIN  LHCENT    LHLN unknown 
   #    4      15     479    2425 
```

```{r LHneurons, echo=FALSE}
LHtype  <- m[grepl("LH", m$cell_type, ignore.case = TRUE), ]
LHLN <- m[grepl("LHLN", m$cell_class, ignore.case = TRUE), ]
LHff_m = subset(m, m$root_id %in% LHff_mids)
LHneurons = unique(bind_rows(LHtype,LHLN,LHff_m))

# get all of a given typed cell type
LHneuronstypes = unique(LHneurons$cell_type)
LHneuronstypes = na.omit(LHneuronstypes)
mLHcelltype =  subset(m, m$cell_type %in% LHneuronstypes)
LHneurons = unique(bind_rows(LHneurons,mLHcelltype))
nrow(LHneurons)
# 2805 neurons

LHneurons  <- LHneurons %>%
  mutate(cell_type = ifelse(grepl("^\\d+$", cell_type) & cell_class == "LHLN", paste0("LHLN", cell_type), cell_type))
LHneurons <- LHneurons %>%
  mutate_all(~ifelse(is.na(.), "unknown", .))
table(LHneurons$cell_class)
   # ALIN  LHCENT    LHLN unknown 
   #    2      42     479    2282 
table(LHneurons$cell_type)

# remove neurons that have not been matched at all (current type = unknown)
unknown1 = subset(LHneurons, LHneurons$cell_type == "unknown")
LHneurons = subset(LHneurons, !LHneurons$cell_type == "unknown")
```
some of the unknown cell type have a number in cf_meta for their group add this to the current meta data of LHneurons
```{r LHneurons, echo=FALSE}
unknown = cf_meta(cf_ids(flywire=unknown1$root_id)) 
unknowntype = subset(unknown, select = c(supervoxel_id, type))
names(unknowntype) = c("supervoxel_id", "cell_type")
unknown1$cell_type<-NULL
unknown1 = left_join(unknown1,unknowntype)


LHneurons = bind_rows(LHneurons, unknown1)
LHneurons <- LHneurons %>%
  mutate_all(~ifelse(is.na(.), "unknown", .))
# 315 unknown that are not LHLNs
LHneurons$cell_type[LHneurons$cell_type == "unknown"] = paste0("unknown",LHneurons$root_id, sep = "")


# summary: 
# 2805 neurons
#    ALIN  LHCENT    LHLN unknown 
      # 2      42     479    2282 
saveRDS(LHneurons, file = "LHneurons.RDS")
```



look again at the feather file for these neurons
```{r ff for LH neurons, echo=FALSE}
# take root_ids from that specific timepoint
# LHneurons$root_id = flywire_updateids(LHneurons$root_id, timestamp = 756)
LHneuronsff = subset(ff, ff$post_pt_root_id %in% LHneurons$root_id)
length(unique(LHneuronsff$post_pt_root_id))
#  2302
# not all !
saveRDS(LHneuronsff, file = "LHneuronsff.RDS")
```
plot
```{r ff for LH neurons, echo=FALSE}
plotLHneuronsff = subset(LHneuronsff, LHneuronsff$neuropil %in% c("LH_L", "LH_R"))
plotLHneuronsff = aggregate(plotLHneuronsff$percent, by = list(plotLHneuronsff$post_pt_root_id), FUN = sum)
plotLHneuronsff %>%
  arrange(Group.1) %>%
  ggplot(aes(x = reorder(Group.1, -x), y = x)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Percentages in LH_R and LH_L",
       x = "post_pt_root_id",
       y = "Percentage") +
  theme_minimal()
```
